---
title: "Reproducible Research: Peer Assessment 1"
author: "Martin Slíva"
date: "2018-02-28"
output: 
  html_document:
    keep_md: true
    
---


## Loading and preprocessing the data

```{r loading_file}

if(!file.exists('activity.csv')){ unzip('activity.zip')}
activity<-read.csv("activity.csv") 


```

Preprocessing of data is not applied.


## What is mean total number of steps taken per day?



```{r histogram}

steps_sum<-with(activity, tapply(steps, date, FUN = sum, na.rm=T)) ### sum steps made every day

hist(steps_sum, breaks = 20, main="Total number of steps taken per day", xlab = "Number of steps", ylab = "Number of days" )


```


```{r mean_median}

steps_mean<- mean(steps_sum)

steps_median<- median(steps_sum)

```

The mean of steps per day is `r steps_mean` and median is `r steps_median`.



## What is the average daily activity pattern?



```{r   maximal_steps}

steps_interval<-with(activity, aggregate(steps, list(interval),FUN=mean, na.rm=T )) ### 
names(steps_interval)<-c("Interval","Steps")
plot(steps_interval$Interval,steps_interval$Steps, type = "l", xlab = "Interval  - hhmm", ylab = "Average number of steps", main = "Average number of steps accross 5 minute intervals")



steps_interval_max<-steps_interval$Interval[which.max(steps_interval$Steps)]


```

The 5-minute interval with maximal average steps is `r steps_interval_max`.


## Imputing missing values

The strategy for inputing missing values: inputing median of 5-minute time interval. 


```{r number_na}

number_na<-sum(is.na(activity))

```

Number of missing values is `r number_na`.


```{r inputing_missing_values}

steps_interval_median<-with(activity, tapply(steps, interval,FUN=median, na.rm=T ))     ### calculation of median for all interval - NA excluded
steps_na<-which(is.na(activity$steps))                                                  ### finding which rows have NA in steps column
activity_new<-activity                                                                  ### creating new variable
for(i in steps_na) {activity_new$steps[i]<-steps_interval_median[[1+activity$interval[i]%/%100*12+activity$interval[i]%%100/5]]}  ### replacement the NA by calculated median


```

```{r histogram_after_inputing_values}


steps_sum_new<-with(activity_new, tapply(steps, date, FUN = sum, na.rm=T))          ### Calculation sums of steps per date  

hist(steps_sum_new, breaks = 20, main="Total number of steps taken per day after missing value inputed ", xlab = "Number of steps", ylab = "Number of days" )


```



```{r mean_new}

 steps_mean_new<-mean(steps_sum_new)

 mean_difference<-steps_mean_new-steps_mean

```


```{r median_new}

 steps_median_new<-median(steps_sum_new)

```


After filling missing value by median of time interval new mean is `r steps_mean_new` and new median is `r steps_median_new`.

The median stayed the same as input of median do not have impact on it. Average changed by `r mean_difference`.


## Are there differences in activity patterns between weekdays and weekends?



```{r data_prepration_for_final_chart}

activity_new$weekend<-(weekdays(as.Date(activity_new$date)) ==weekdays(as.Date("2018-02-25")) | weekdays(as.Date(activity_new$date)) ==weekdays(as.Date("2018-02-24")))              ### creating variable for division between weekdays and weekends, particular dates for Sunday and Saturday are taken to avoid problems with Locales

steps_interval_new<-with(activity_new, aggregate(steps, by=list(interval,weekend),FUN=mean))

steps_interval_new$day_type[steps_interval_new$Group.2]<-"Weekend"
steps_interval_new$day_type[!steps_interval_new$Group.2]<-"Weekday"
names(steps_interval_new)[1:3]<-c("Time_interval", "Weekend", "Average_steps")



```




```{r final_chart }
library(ggplot2)

final_chart <- ggplot(steps_interval_new, aes(x=Time_interval, y=Average_steps))+geom_line(colour="blue")+facet_grid(day_type~.)+ggtitle("Average steps taken during weekdays and weekends")

final_chart




```

